######################################################################### 
#-----------------------------------------------------------------------#
# Created on:    2022/11/8                                              #
# Updated on:    2022/12/10                                             #
# EMAIL:         dengfy@umich.edu                                       #
#                Umich  Feiyang Deng.                                   #
#########################################################################

##### Data Analysis Script #####
```{r}
rm(list=ls())
setwd("E:/Biostat Study/BIOSTAT 650/Group Project")
set.seed(123)

library('readxl')
library('ggplot2')
library('ggpubr')
library('tidyverse')
library('mice')
library('VIM')
library('MASS')
library('car')
library("DMwR2")
library("Hmisc")
library("dplyr")

source('code/utils.R')
```

#### import data ####
```{r}
setwd("E:/Biostat Study/BIOSTAT 650/Group Project")
inputFile <- "Depression Data.xls"
raw_data <- read_excel(inputFile, sheet=1, na='NA')
raw_data <- as.data.frame(raw_data)

summary(raw_data)
```

#### Visualize missing values
```{r}
md.pattern(raw_data)
aggr(raw_data, prop=F, numbers=T)

```
###############################################################
#### Descriptive analysis for complete and incomplete data ####
###############################################################


## add indicator to identify complete rows
```{r}
na_rows <- which(rowSums(is.na(raw_data)) > 0)
raw_data$complete <- 1
raw_data$complete[na_rows] <- 0
raw_data$complete <- as.factor(raw_data$complete)

```

### Descriptive analysis of AGE4, TOTMMSE4, TOTIADL4, TOTADL4, EE46
## All continuous variables (except outcome) and one categorical predictor of interest
#### Boxplot  ####
```{r}

ggplot(raw_data, aes(x=complete, y=AGE4, fill=complete)) +
  geom_boxplot()
ggplot(raw_data, aes(x=complete, y=TOTMMSE4, fill=complete)) +
  geom_boxplot()
ggplot(raw_data, aes(x=complete, y=TOTIADL4, fill=complete)) +
  geom_boxplot()
ggplot(raw_data, aes(x=complete, y=TOTADL4, fill=complete)) +
  geom_boxplot()
ggplot(raw_data, aes(x=complete, y=EE46, fill=complete)) +
  geom_boxplot()
```

#### Density ####
```{r}
ggplot(raw_data, aes(x=AGE4, fill=complete)) +
  geom_density(alpha=.25)
ggplot(raw_data, aes(x=TOTMMSE4, fill=complete)) +
  geom_density(alpha=.25)
ggplot(raw_data, aes(x=TOTIADL4, fill=complete)) +
  geom_density(alpha=.25)
ggplot(raw_data, aes(x=TOTADL4, fill=complete)) +
  geom_density(alpha=.25)
```

#### Histogram ####
```{r}
ggplot() +
  geom_histogram(data=raw_data[-na_rows,], 
                 binwidth = 0.5,
                 aes(x=EE46, y=..count../(1682-length(na_rows))),
                 fill = 13, alpha = 0.7) +
  geom_histogram(data=raw_data[na_rows,], 
                 binwidth = 0.5,
                 aes(x=EE46, y=-..count../sum(1-is.na(raw_data$EE46[na_rows]))),
                 fill = 10, alpha = 0.7) +
  ylab('Incomplete data VS Complete data') +
  ggtitle('Histogram of EE46') +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate(geom = "label", 
           x = 4, y = 0.4, 
           label = "EE46 of Complete data",
           fill = 12, 
           color = "white", hjust = 1) +
  annotate(geom = "label", 
           x = 4, y = -0.3, 
           label = "EE46 of Inomplete data",
           fill = 14, 
           color = "white", hjust = 1)
```

## test if same distribution AGE4, TOTMMSE4, TOTIADL4, TOTADL4, EE46
```{r}
ks.test(raw_data$AGE4[-na_rows], raw_data$AGE4[na_rows], na.omit=TRUE)
ks.test(raw_data$TOTIADL4[-na_rows], raw_data$TOTIADL4[na_rows], na.omit=TRUE)
ks.test(raw_data$TOTADL4[-na_rows], raw_data$TOTADL4[na_rows], na.omit=TRUE)
ks.test(raw_data$TOTMMSE4[-na_rows], raw_data$TOTMMSE4[na_rows], na.omit=TRUE)
ks.test(raw_data$NKIDS4[-na_rows], raw_data$NKIDS4[na_rows], na.omit=TRUE)
chisq.test(raw_data$EE46, raw_data$complete)
```

############################################
```{r}
aggregate(raw_data$NKIDS4, by=list(raw_data$complete), mean, na.rm=TRUE)
```








####################
#### Imputation ####
####################
# column index
# Categorical variables : GRADE, USBORN, marstat, nkids4, health4, khyper41, mdiab41, nfrac41, u43s, cc43, ee46, hha4, oo49lang, Male
# indeces : GRADE(2), USBORN(3), MARSTAT4(5), NKIDS4(6), HEALTH4(7), KHYPER41(8),
```{r}
contin_idx <- c(4,12,14,15)
outcome_idx <- c(13)
cate_na_idx <- c(2,6,8,9,10,11,16,17,18)# index of categorical with missing values
cate_full_idx <- c(3,5,7,19,20)
other_idx <- c(1,21)
colnames(raw_data[,cate_full_idx])

summary(raw_data)
```

################## Imputation  #########################
```{r}
# raw_data$GRADE <- as.factor(raw_data$GRADE)
raw_data$USBORN <- as.factor(raw_data$USBORN)
# raw_data$AGE4 <- as.factor(raw_data$AGE4)
raw_data$MARSTAT4 <- as.factor(raw_data$MARSTAT4)
# raw_data$NKIDS4 <- as.factor(raw_data$NKIDS4)
raw_data$HEALTH4 <- ordered(raw_data$HEALTH4) ## order
raw_data$KHYPER41 <- as.factor(raw_data$KHYPER41)
raw_data$MDIAB41 <- as.factor(raw_data$MDIAB41)
raw_data$NFRAC41 <- as.factor(raw_data$NFRAC41)
raw_data$U43S <- as.factor(raw_data$U43S)
raw_data$CC43 <- ordered(raw_data$CC43) ## order
raw_data$EE46 <- ordered(raw_data$EE46) ## order
raw_data$HHA4 <- as.factor(raw_data$HHA4)
raw_data$OO49LANG <- as.factor(raw_data$OO49LANG)
raw_data$MALE <- as.factor(raw_data$MALE)
outcome_na_idx <- which(is.na(raw_data$CESDTOT4))
colnames(raw_data[-c(1,21)])
mice_tmp <- mice(raw_data[-c(1,21)],m=5,maxit=10,seed=500) ## 4 best result so far
mice_data <- complete(mice_tmp,4)[-outcome_na_idx,]
# mice_data <- complete(mice_tmp,5)
summary(mice_data)
mice_data$HEALTH4 <- factor(mice_data$HEALTH4, ordered = FALSE) ## order
mice_data$CC43 <- factor(mice_data$CC43, ordered = FALSE) ## order
mice_data$EE46 <- factor(mice_data$EE46, ordered = FALSE) ## order
new_data <- mice_data
table(mice_data$GRADE)
table(mice_data$NKIDS4)
t <- lm(CESDTOT4~.,data=mice_data)
summary(t)
par(mfrow=c(2,2)) 
plot(t)  ## double bow  transform arcsin(sqrt(y))
par(mfrow=c(1,1))

# mice_data$CESDTOT4 <- sqrt(mice_data$CESDTOT4)
# t <- lm(CESDTOT4~.,data=mice_data)
# summary(t)
# par(mfrow=c(2,2)) 
# plot(t)  ## double bow  transform arcsin(sqrt(y))
# par(mfrow=c(1,1))
```
######################################################




########################################


############################################
#### Descriptive analysis for new data ####
###########################################
# select(CESDTOT4,USBORN,MARSTAT4,HEALTH4,OO49LANG,MALE,GRADE,
#        NKIDS4,KHYPER41,MDIAB41,NFRAC41,U43S,CC43,EE46,HHA4,AGE4,
#        TOTMMSE4,TOTIADL4,TOTADL4) %>%
```{r}
colnames(new_data)
library("gtsummary")
new_data %>%
  tbl_summary(missing = "no",
              type = list(CESDTOT4 ~ 'continuous',
                          USBORN ~ 'categorical',
                          MARSTAT4 ~ 'categorical',
                          HEALTH4 ~ 'categorical',
                          OO49LANG ~ 'categorical',
                          MALE ~ 'categorical',
                          GRADE ~ 'continuous',
                          NKIDS4 ~ 'continuous',
                          KHYPER41 ~ 'categorical',
                          MDIAB41 ~ 'categorical',
                          NFRAC41 ~ 'categorical',
                          U43S ~ 'categorical',
                          CC43 ~ 'categorical',
                          EE46 ~ 'categorical',
                          HHA4 ~ 'categorical',
                          AGE4 ~ 'continuous',
                          TOTMMSE4 ~ 'continuous',
                          TOTIADL4 ~ 'continuous',
                          TOTADL4 ~ 'categorical'),
              statistic = list(all_continuous() ~ "{mean} ({sd}) min:{min} max:{max} IQR:[{p25}, {p75}]",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(2,2),
                            all_categorical() ~ c(0,2))
  )

```

#### boxplot ####
```{r}
p1 <- ggplot(new_data, aes(y=CESDTOT4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p2 <- ggplot(new_data, aes(y=TOTIADL4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p3 <- ggplot(new_data, aes(y=TOTADL4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p4 <- ggplot(new_data, aes(y=AGE4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p5 <- ggplot(new_data, aes(y=TOTMMSE4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p6 <- ggplot(new_data, aes(y=GRADE)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")

ggarrange(p1,p2,p3,p4,p5,p6,
          ncol = 3, nrow = 2)

```

##### density of outcome and continuous variables ####
```{r}
ggplot(new_data, aes(x=CESDTOT4)) +
  geom_density(alpha=.5, fill="lightblue", color="blue")

ggplot(new_data, aes(x=TOTIADL4)) +
  geom_density(alpha=.5, fill="lightblue", color="blue")
ggplot(new_data, aes(x=TOTADL4)) +
  geom_density(alpha=.5, fill="lightblue", color="blue")
ggplot(new_data, aes(x=AGE4)) +
  geom_density(alpha=.5, fill="lightblue", color="blue")
ggplot(new_data, aes(x=TOTMMSE4)) +
  geom_density(alpha=.5, fill="lightblue", color="blue")
```

##### histogram of ordical variables ####
```{r}
ggplot(new_data, aes(x=as.factor(GRADE))) +
  geom_histogram(color="lightblue",stat="count", fill="#69b3a2", alpha=0.7)
ggplot(new_data, aes(x=as.factor(HEALTH4))) +
  geom_histogram(color="lightblue",stat="count", fill="#69b3a2", alpha=0.7)
ggplot(new_data, aes(x=as.factor(TOTIADL4))) +
  geom_histogram(color="lightblue",stat="count", fill="#69b3a2", alpha=0.7)
ggplot(new_data, aes(x=as.factor(EE46))) +
  geom_histogram(color="lightblue",stat="count", fill="#69b3a2", alpha=0.7)

```

##### piechart of categorical variables ####
```{r}
p <- ggplot(data=new_data,mapping = aes(x='Content',y=as.factor(EE46),fill=as.factor(EE46)))+ 
  geom_bar(stat = 'identity', position = 'stack', width = 1)
p + coord_polar(theta = 'y') + labs(x = 'Category', y = 'number', title = 'pie') + 
  theme(axis.text = element_blank())

p <- ggplot(data=new_data,mapping = aes(x='Content',y=as.factor(MARSTAT4),fill=as.factor(MARSTAT4)))+ 
  geom_bar(stat = 'identity', position = 'stack', width = 1)
p + coord_polar(theta = 'y') + labs(x = 'Category', y = 'number', title = 'pie') + 
  theme(axis.text = element_blank())
```

# psych::pairs.panels(new_data)

##### outcome vs Xi ####
```{r}
ggplot(new_data, aes(x = TOTIADL4, y = CESDTOT4)) + 
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)
ggplot(new_data, aes(x = TOTADL4, y = CESDTOT4)) + 
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)
ggplot(new_data, aes(x = AGE4, y = CESDTOT4)) + 
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)
ggplot(new_data, aes(x = TOTMMSE4, y = CESDTOT4)) + 
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)

```

##### outcome vs categorical ####
```{r}
ggplot(new_data, aes(x = EE46,y=CESDTOT4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = as.factor(GRADE), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = as.factor(NKIDS4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = HEALTH4, y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = as.factor(AGE4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = MARSTAT4, y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")

ggplot(new_data, aes(x = as.factor(TOTADL4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = as.factor(TOTIADL4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggplot(new_data, aes(x = as.factor(TOTMMSE4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")




p1 <- ggplot(new_data, aes(x = as.factor(TOTIADL4), y = CESDTOT4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
p2 <- ggplot(new_data, aes(x = EE46,y=CESDTOT4)) +
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray")
ggarrange(p1,p2,
          ncol = 2, nrow = 1)


```

#### correlation of covariates ####
## just for reference
```{r}
library("corrplot")
summary(new_data)
contin_idx_cor <- c(1,3,5,11,13,14)
contin_data <- new_data[,contin_idx_cor]
cate_data <- new_data[,-contin_idx_cor]
for(i in 1:length(cate_data)){
  cate_data[,i] <- as.numeric(as.character(cate_data[,i]))
}
summary(cate_data)

cormat = cor(contin_data, method = "pearson")
pres <- cor.mtest(contin_data, conf.level = .95)
corrplot.mixed(cormat, lower.col = "black", number.cex = 1,p.mat = pres$p, sig.level = .05)

cormat = cor(cate_data, method = "spearman")
pres <- cor.mtest(cate_data, conf.level = .95)
corrplot.mixed(cormat, lower.col = "black", number.cex = 1,p.mat = pres$p, sig.level = .05)

cor.test(contin_data$TOTIADL4, contin_data$TOTADL4)

table(contin_data$TOTADL4)
```
######################################


########################################################
#### Bivariate analysis (outcome vs. Xi regression) ####
########################################################
# Questions: 1. how to deal with categorical variables with categories > 5 ??  eg. GRADE KIDS
#            2. do we need to make group for continuous variables ??  eg. AGE IDAL
#            3. in data there are to major groups (physical health variables and mental health variables)
#               is there any relationship inside or between physical health variables and mental health variables ??
```{r}
head(new_data)
#2 grade not significant
M2=lm(CESDTOT4~GRADE,data=new_data)
summary(M2)
# not significant
BM2=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+GRADE,data=new_data)
summary(BM2)
```

## after transform grade could be significant
```{r}
table(new_data$GRADE)
new_data$GRADE_Cat <- 0
new_data$GRADE_Cat[new_data$GRADE>=6 & new_data$GRADE<=11] <- 1
new_data$GRADE_Cat[new_data$GRADE>=12] <- 2
new_data$GRADE_Cat <- as.factor(new_data$GRADE_Cat)
table(new_data$GRADE_Cat)

M2=lm(CESDTOT4~GRADE_Cat,data=new_data)
summary(M2)
anova(M2)
BM2=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+as.factor(GRADE_Cat),data=new_data)
anova(BM2)


#3 USBORN  significant
M3=lm(CESDTOT4~USBORN,data=new_data)
summary(M3)
# significant
BM3=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+USBORN,data=new_data)
summary(BM3)

#4 AGE4  significant
M4=lm(CESDTOT4~AGE4,data=new_data)
summary(M4)
#not significant
BM4=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+AGE4,data=new_data)
summary(BM4)

table(new_data$AGE4)
new_data$AGE4_Cat <- 0
new_data$AGE4_Cat[new_data$AGE4>=80 & new_data$AGE4<=89] <- 1
new_data$AGE4_Cat[new_data$AGE4>=90] <- 2
new_data$AGE4_Cat <- as.factor(new_data$AGE4_Cat)

# new_data$AGE4_Cat <- split_IQR(new_data$AGE4)
#significant after cate
table(new_data$AGE4_Cat)

M4=lm(CESDTOT4~AGE4_Cat,data=new_data)
summary(M4)
anova(M4)
BM4=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+AGE4_Cat,data=new_data)
summary(BM4)

#5 MARSTAT4 significant
new_data$MARSTAT4 <- as.factor(new_data$MARSTAT4)
M5=lm(CESDTOT4~MARSTAT4,data=new_data)
summary(M5)
anova(M5)
# significant
BM5=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+MARSTAT4,data=new_data)
anova(BM5)

table(new_data$MARSTAT4)
new_data$MARSTAT4_Cat <- 0
new_data$MARSTAT4_Cat[new_data$MARSTAT4==3] <- 1
table(new_data$MARSTAT4_Cat)
new_data$MARSTAT4_Cat <- as.factor(new_data$MARSTAT4_Cat)

M5=lm(CESDTOT4~MARSTAT4_Cat,data=new_data)
summary(M5)

BM5=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+MARSTAT4_Cat,data=new_data)
anova(BM5)

#6 NKIDS4 not significant
M6=lm(CESDTOT4~NKIDS4,data=new_data)
summary(M6)
# not significant
BM6=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+NKIDS4,data=new_data)
summary(BM6)

table(new_data$NKIDS4)
new_data$NKIDS4_Cat <- 0
new_data$NKIDS4_Cat[new_data$NKIDS4>=4] <- 1
new_data$NKIDS4_Cat <- as.factor(new_data$NKIDS4_Cat)
table(new_data$NKIDS4_Cat)

M6=lm(CESDTOT4~as.factor(NKIDS4_Cat),data=new_data)
summary(M6)
BM6=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+as.factor(NKIDS4_Cat),data=new_data)
summary(BM6)

#7 HEALTH4 significant
M7=lm(CESDTOT4~factor(HEALTH4),data=new_data)
summary(M7)
anova(M7)
# significant
BM7=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+factor(HEALTH4),data=new_data)
anova(BM7)

table(new_data$HEALTH4)
new_data$HEALTH4_Cat <- new_data$HEALTH4
new_data$HEALTH4_Cat[new_data$HEALTH4==2] <- 1
table(new_data$HEALTH4_Cat)
M7=lm(CESDTOT4~factor(HEALTH4_Cat),data=new_data)
summary(M7)
BM7=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+factor(HEALTH4_Cat),data=new_data)
anova(BM7)

#8 KHYPER41 significant
M8=lm(CESDTOT4~KHYPER41,data=new_data)
summary(M8)
# significant
BM8=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+KHYPER41,data=new_data)
summary(BM8)

#9 MDIAB41 significant
M9=lm(CESDTOT4~MDIAB41,data=new_data)
summary(M9)
# not significant
BM9=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+MDIAB41,data=new_data)
summary(BM9)

#10 NFRAC41 not significant
M10=lm(CESDTOT4~NFRAC41,data=new_data)
summary(M10)

# not significant
BM10=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+NFRAC41,data=new_data)
summary(BM10)

#11 U43S significant
M11=lm(CESDTOT4~U43S,data=new_data)
summary(M11)
# significant
BM11=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+U43S,data=new_data)
summary(BM11)


#12 TOTMMSE4 significant
M12=lm(CESDTOT4~TOTMMSE4,data=new_data)
summary(M12)
# significant
BM12=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+TOTMMSE4,data=new_data)
summary(BM12)
```
## transform is not significant after stepwise
```{r}
table(new_data$TOTMMSE4)
new_data$TOTMMSE4_Cat <- 0
new_data$TOTMMSE4_Cat[new_data$TOTMMSE4>=17 & new_data$TOTMMSE4<=24] <- 1
new_data$TOTMMSE4_Cat[new_data$TOTMMSE4>=25] <- 2
new_data$TOTMMSE4_Cat <- as.factor(new_data$TOTMMSE4_Cat)
table(new_data$TOTMMSE4_Cat)

M12=lm(CESDTOT4~TOTMMSE4_Cat,data=new_data)
summary(M12)
BM12=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+TOTMMSE4_Cat,data=new_data)
summary(BM12)

```
#14 TOTIADL4 significant
```{r}
M14=lm(CESDTOT4~TOTIADL4,data=new_data)
summary(M14)

table(new_data$TOTIADL4)
new_data$TOTIADL4_Cat <- 0
new_data$TOTIADL4_Cat[new_data$TOTIADL4>=5 & new_data$TOTIADL4<=8] <- 1
new_data$TOTIADL4_Cat[new_data$TOTIADL4>=9] <- 2
new_data$TOTIADL4_Cat <- as.factor(new_data$TOTIADL4_Cat)
table(new_data$TOTIADL4_Cat)
M15=lm(CESDTOT4~TOTIADL4_Cat,data=new_data)
summary(M15)


#15 totadl4 significant
M15=lm(CESDTOT4~TOTADL4,data=new_data)
summary(M15)
# significant
BM15=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+TOTADL4,data=new_data)
summary(BM15)

table(new_data$TOTADL4)
new_data$TOTADL4_Cat <- 0
new_data$TOTADL4_Cat[new_data$TOTADL4!=0] <- 1
new_data$TOTADL4_Cat <- as.factor(new_data$TOTADL4_Cat)

M15=lm(CESDTOT4~TOTADL4,data=new_data)
summary(M15)
BM15=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+TOTADL4_Cat,data=new_data)
summary(BM15)

#16 CC43 significant
M16=lm(CESDTOT4~CC43,data=new_data)
summary(M16)
# significant
BM16=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+CC43,data=new_data)
summary(BM16)

#17 EE46 significant
M17=lm(CESDTOT4~EE46,data=new_data)
summary(M17)
anova(M17)

table(new_data$EE46)
new_data$EE46_Cat <- new_data$EE46
new_data$EE46_Cat[new_data$EE46==4] <- 1
M17=lm(CESDTOT4~EE46_Cat,data=new_data)
summary(M17)

#18 hha4 -significant
M18=lm(CESDTOT4~HHA4,data=new_data)
summary(M18)
#not significant 
BM18=lm(CESDTOT4~TOTIADL4+as.factor(EE46)+HHA4,data=new_data)
summary(BM18)

#19 oo49lang -significant
M19=lm(CESDTOT4~OO49LANG,data=new_data)
summary(M19)
# significant
BM19<-lm(CESDTOT4~TOTIADL4+as.factor(EE46)+OO49LANG,data=new_data)
summary(BM19)

#20 Male -significant
M20=lm(CESDTOT4~MALE,data=new_data)
summary(M20)
# siginificant
BM20<-lm(CESDTOT4~TOTIADL4+as.factor(EE46)+MALE,data=new_data)
summary(BM20)

```

###############################
```{r}
colnames(new_data)
new_data %>%
  tbl_summary(missing = "no",
              type = list(NKIDS4 ~ 'continuous',
                          TOTMMSE4 ~ 'continuous',
                          TOTIADL4 ~ 'continuous',
                          TOTADL4 ~ 'continuous'),
              statistic = list(NKIDS4 ~ "{mean} ({sd})",
                               TOTMMSE4 ~ "{mean} ({sd})",
                               TOTIADL4 ~ "{mean} ({sd})",
                               TOTADL4 ~ "{mean} ({sd})"),
              digits = list(all_continuous() ~ c(3,3))
  )



```
##########################
##### Building Models ####
##########################
```{r}
colnames(new_data)

demo_idx <- c(20,2,21,4,18,19) # Grade_Cat USborn Age_Cat Marstat language Male
physi_idx <- c(7,8,9,10,17) # 
cong_idx <- c(11)
outcome_idx <- c(12)
poi_idx <- c(13,16)

for (i in physi_idx){
  new_data[,i] <- as.numeric(as.character(new_data[,i]))
}

new_data$Chronic <- new_data$KHYPER41+new_data$MDIAB41+new_data$NFRAC41+
  new_data$U43S+new_data$HHA4

for (i in physi_idx){
  new_data[,i] <- as.factor(new_data[,i])
}

colnames(new_data)
table(new_data$Chronic)
new_data$Chronic[new_data$Chronic==4] <- 3
new_data$Chronic <- as.factor(new_data$Chronic)
table(new_data$Chronic)

physi_idx <- c(29)


select_data <- new_data[,c(demo_idx,physi_idx,cong_idx,outcome_idx,poi_idx)]
colnames(select_data)
summary(select_data)
```
#### Full model ####
```{r}
full.model <- lm(CESDTOT4~., data = select_data)
summary(full.model)
```
### plots
```{r}
par(mfrow=c(2,2)) 
plot(full.model)  ## sqrt(y)
par(mfrow=c(1,1))
```

### partial regression plot
```{r}
avPlots(full.model)
```
### Residual plots
```{r}
residualPlots(full.model,type="response")
```

##### Transform ####
```{r}
select_data$CESDTOT4 <- sqrt(select_data$CESDTOT4)
full.model.t <- lm(CESDTOT4~., data = select_data)
summary(full.model.t)
```
### plots
```{r}
par(mfrow=c(2,2)) 
plot(full.model.t)
par(mfrow=c(1,1))
```
### partial regression plot
```{r}
avPlots(full.model.t)
```
### Residual plots
```{r}
residualPlots(full.model.t,type="response")
```

### full model doesn't have multicollinearity
```{r}
vif_values <- vif(full.model.t)[,1]
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "Orchid")
abline(v = 5, lwd = 3, lty = 2)

```
#### Outliers ####
```{r}
full_yhat = full.model.t$fitted.values
full_res = full.model.t$residuals #m1.yhat-fev
full_h = hatvalues(full.model.t) #leverage
full_r = rstandard(full.model.t) #internally studentized residuals
full_rr = rstudent(full.model.t) #externally studentized residuals

n = nrow(select_data)
p = full.model.t$rank # dimensions
full_sigma = sqrt( sum(full_res^2)/(n-p) )
full_z = full_res/full_sigma #standardized residual

outlier <- rep(0, length(full_z))
outlier[which(abs(full_rr) >= 3)] <- 1
sum(outlier)
outlier <- as.factor(outlier)

qplot(x=full_yhat, y=full_rr, col=outlier)
qplot(x=new_data$TOTIADL4, y=new_data$CESDTOT4, col=outlier)

tmp <- new_data[which(abs(full_rr) >= 3),]

full.model.t2 <- lm(CESDTOT4~., data = select_data[-which(abs(full_rr) >= 3),])
summary(full.model.t2)
```
### plots
```{r}
par(mfrow=c(2,2)) 
plot(full.model.t2)
par(mfrow=c(1,1))

```
#### difference after taking out outliers
```{r}
(coefficients(full.model.t) - coefficients(full.model.t2))/coefficients(full.model.t)*100
```

####################################################





###################
#### Diagnosis ####
###################
```{r}
main.model <- full.model.t
```
### plots
```{r}
par(mfrow=c(2,2)) 
plot(main.model)
par(mfrow=c(1,1))
```
### partial regression plot
```{r}
avPlots(main.model)
```
### Residual plots
```{r}
residualPlots(main.model,type="response")
```
### multicollinearity
```{r}
vif_values <- vif(main.model)[,1]
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "Orchid")
abline(v = 10, lwd = 3, lty = 2)
ggplot(select_data, aes(x = TOTIADL4, y = CESDTOT4)) + 
  geom_point(color = "blue") +
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)
main_yhat = main.model$fitted.values
main_res = main.model$residuals #m1.yhat-fev
main_h = hatvalues(main.model) #leverage
main_r = rstandard(main.model) #internally studentized residuals
main_rr = rstudent(main.model) #externally studentized residuals
n = nrow(select_data)
p = main.model$rank # dimensions
main_sigma = sqrt( sum(main_res^2)/(n-p) )
main_z = main_res/main_sigma #standardized residual
```
## outlier
```{r}
outlier <- rep(0, length(main_z))
outlier[which(abs(main_rr) >= 3)] <- 1
sum(outlier)
outlier <- as.factor(outlier)

qplot(x=main_yhat, y=main_rr, col=outlier)

```
### normal
```{r}
qqPlot(main_z)
hist(main_z)

shapiro.test(main_z) ## reject but only for small sample size (<50)
ks.test(main_z, "pnorm")
library('nortest')
lillie.test(main_z)

```

### empirical cdf of main_z and N(0,1)
```{r}
sample_plot <- data.frame(z=main_z, rn=rnorm(length(main_z)))
ggplot(sample_plot) +
  stat_ecdf(aes(main_z),color="red", size=1, alpha=0.7) +
  stat_ecdf(aes(rn),color="blue", size=1, alpha=0.7) +
  ylab('') + xlab('') +
  ggtitle('Empirical cdf of standardized residual and N(0,1)') +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate(geom = "label", 
           x = 4, y = 0.6, 
           label = "cdf of standardized residual",
           fill = "red", 
           color = "white", hjust = 1) +
  annotate(geom = "label", 
           x = 3, y = 0.4, 
           label = "cdf of N(0, 1)",
           fill = "blue", 
           color = "white", hjust = 1)
```
### independence
```{r}
main_res1 <- main_res[1:length(main_res)-1]
main_res2 <- main_res[2:length(main_res)]
qplot(x=main_res1, y=main_res2)
```
################
# Transform ####
################
```{r}
summary(main.model)
```
## Scale
```{r}
select_data$TOTIADL4 <- scale(select_data$TOTIADL4, scale = FALSE)
select_data$TOTMMSE4 <- scale(select_data$TOTMMSE4, scale = FALSE)
main.model2 <- lm(formula = CESDTOT4 ~ .+I(TOTIADL4^2), data = select_data)
summary(main.model2)
```
### plots
```{r}
par(mfrow=c(2,2))
plot(main.model2)
par(mfrow=c(1,1))

```
### partial regression plot
```{r}
avPlots(main.model2)
```
### Residual plots
```{r}
residualPlots(main.model2,type="response")
```
### multicollinearity
```{r}
vif_values <- vif(main.model2)[,1]
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "Orchid")
abline(v = 10, lwd = 3, lty = 2)
ggplot(select_data, aes(x = TOTIADL4, y = CESDTOT4)) + 
  geom_point(color = "blue") +
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE)

main_yhat2 = main.model2$fitted.values
main_res2 = main.model2$residuals #m1.yhat-fev
main_h2 = hatvalues(main.model2) #leverage
main_r2 = rstandard(main.model2) #internally studentized residuals
main_rr2 = rstudent(main.model2) #externally studentized residuals

n = nrow(select_data)
p = main.model2$rank # dimensions
main_sigma2 = sqrt( sum(main_res2^2)/(n-p) )
main_z2 = main_res2/main_sigma2 #standardized residual
```
## outlier
```{r}
outlier <- rep(0, length(main_z2))
outlier[which(abs(main_rr2) >= 3)] <- 1
sum(outlier)
outlier <- as.factor(outlier)

qplot(x=main_yhat2, y=main_rr2, col=outlier,
      xlab="Fitted values", ylab="External studentized residuals")

```
### normal
```{r}
qqPlot(main_z2)
hist(main_z2)

shapiro.test(main_z2) ## reject but only for small sample size (<50)
ks.test(main_z2, "pnorm")
library('nortest')
lillie.test(main_z2)

```
####################################
#### Age_Cat significant _> not ####
####################################
## possible reason: relationship between Age_Cat and IADL
```{r}
kruskal.test(select_data$AGE4_Cat, select_data$TOTIADL4)
ggplot(select_data, aes(x = AGE4_Cat, y = TOTIADL4)) + 
  geom_boxplot(fill="lightgreen", alpha=0.7, color="DarkSlateGray") +
  ggtitle("TOTIADL4 vs. AGE4_Cat")
```
## sequential
```{r}
lr1 <- lm(CESDTOT4~GRADE_Cat+USBORN+AGE4_Cat+MARSTAT4+OO49LANG+MALE, data = select_data)
summary(lr1)
lr2 <- lm(CESDTOT4~GRADE_Cat+USBORN+AGE4_Cat+MARSTAT4+OO49LANG+MALE+
            Chronic+TOTMMSE4, data = select_data)
summary(lr2)
lr3 <- lm(CESDTOT4~GRADE_Cat+USBORN+AGE4_Cat+MARSTAT4+OO49LANG+MALE+
            Chronic+TOTMMSE4+
            EE46, data = select_data)
summary(lr3)
lr4 <- lm(CESDTOT4~GRADE_Cat+USBORN+AGE4_Cat+MARSTAT4+OO49LANG+MALE+
            Chronic+TOTMMSE4+
            +EE46+TOTIADL4+I(TOTIADL4^2), data = select_data)
summary(lr4)

seq_table1 <- summary(lr1)[[4]][,c(1,4)]
coef <- as.data.frame(rownames(seq_table1))
colnames(coef) <- c('coef')
seq_table1 <- cbind(coef, seq_table1)
seq_table1

seq_table2 <- summary(lr2)[[4]][,c(1,4)]
coef <- as.data.frame(rownames(seq_table2))
colnames(coef) <- c('coef')
seq_table2 <- cbind(coef, seq_table2)
seq_table2

seq_table3 <- summary(lr3)[[4]][,c(1,4)]
coef <- as.data.frame(rownames(seq_table3))
colnames(coef) <- c('coef')
seq_table3 <- cbind(coef, seq_table3)
seq_table3

seq_table4 <- summary(lr4)[[4]][,c(1,4)]
coef <- as.data.frame(rownames(seq_table4))
colnames(coef) <- c('coef')
seq_table4 <- cbind(coef, seq_table4)
seq_table4

seq_table <- right_join(right_join(seq_table1, seq_table2, 
                                   by='coef',suffix = c(".1", ".2")),
                        seq_table3, by='coef')
seq_table <- right_join(seq_table, seq_table4, 
                        by='coef')
seq_table
write.csv(seq_table, "E:/Biostat Study/BIOSTAT 650/Group Project/seq_table.csv")
anova(lr1,lr2)
anova(lr2,lr3)
anova(lr3,lr4)

```

#####################
#### Interaction ####
#####################
```{r}
summary(main.model)
summary(main.model2)
write.csv(select_data, "E:/Biostat Study/BIOSTAT 650/Group Project/select_data.csv")
summary(select_data)
main.model2$terms[[3]]
```
#### interaction discovery
```{r}
p1 <- ggplot(select_data[select_data$MARSTAT4==1,], aes(x=TOTIADL4,y=CESDTOT4)) +
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE, size=2) +
  ggtitle("Married")
p2 <- ggplot(select_data[select_data$MARSTAT4==2,], aes(x=TOTIADL4,y=CESDTOT4)) +
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE, size=2) +
  ggtitle("Separated")
p3 <- ggplot(select_data[select_data$MARSTAT4==3,], aes(x=TOTIADL4,y=CESDTOT4)) +
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE, size=2) +
  ggtitle("Divorced")
p4 <- ggplot(select_data[select_data$MARSTAT4==4,], aes(x=TOTIADL4,y=CESDTOT4)) +
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE, size=2) +
  ggtitle("Widowed")
p5 <- ggplot(select_data[select_data$MARSTAT4==5,], aes(x=TOTIADL4,y=CESDTOT4)) +
  geom_point(color = "blue")+
  geom_smooth(method = lm, color = "red", fill="#69b3a2", se = TRUE, size=2) +
  ggtitle("Never Married")
p <- ggarrange(p1,p2,p3,p4,p5,
          ncol = 5, nrow = 1)
annotate_figure(p, top = text_grob("CESDTOT4 vs. TOTIADL4 of Different Marital Status",
                                   face = "bold", size = 14))

inter.model <- lm(formula = CESDTOT4 ~ .+I(TOTIADL4^2) + TOTIADL4*MARSTAT4, data = select_data)
summary(inter.model)
anova(inter.model)
```
### plots
```{r}
par(mfrow=c(2,2))
plot(inter.model)
par(mfrow=c(1,1))
```
### partial regression plot
```{r}
avPlots(inter.model)
```
### Residual plots
```{r}
residualPlots(inter.model)
```
### multicollinearity

```{r}
vif_values <- vif(inter.model,type = 'predictor')[,1]
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "Orchid")
abline(v = 10, lwd = 3, lty = 2)

inter_yhat = inter.model$fitted.values
inter_res = inter.model$residuals #m1.yhat-fev
inter_h = hatvalues(inter.model) #leverage
inter_r = rstandard(inter.model) #internally studentized residuals
inter_rr = rstudent(inter.model) #externally studentized residuals

n = nrow(select_data)
p = inter.model$rank # dimensions
inter_sigma = sqrt( sum(inter_res^2)/(n-p) )
inter_z = inter_res/inter_sigma #standardized residual
```

## outlier
```{r}
outlier <- rep(0, length(inter_z))
outlier[which(abs(inter_rr) >= 3)] <- 1
sum(outlier)
outlier <- as.factor(outlier)
tmp <- new_data[which(abs(inter_rr) >= 3),]

qplot(x=inter_yhat, y=inter_rr, col=outlier,
      xlab="Fitted values", ylab="External studentized residuals", 
      main="External Studentized Residule vs. Fitted Values")


cooksd <- cooks.distance(inter.model)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance") 
abline(h = 0.01, col="red")  
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>0.01,names(cooksd),""), col="red")  

```

### normal
```{r}
qqPlot(inter_z, main='QQplot of Standardized Residuals')
hist(inter_z)

shapiro.test(inter_z) ## reject but only for small sample size (<50)
ks.test(inter_z, "pnorm")
library('nortest')
lillie.test(inter_z)
```
### independence
```{r}
inter_res1 <- inter_res[1:length(inter_res)-1]
inter_res2 <- inter_res[2:length(inter_res)]
qplot(x=inter_res1, y=inter_res2,col='red',xlab='res i-i', ylab='res i', 
      main='residual i vs. residual i-1')

tmp <- summary(inter.model)$coefficients
write.csv(tmp, "E:/Biostat Study/BIOSTAT 650/Group Project/coef_table.csv")

# mean quantile
aggregate(new_data$CESDTOT4, by=list(select_data$AGE4_Cat), mean)
M20=lm(CESDTOT4~Chronic,data=select_data)
summary(M20)
anova(M20)

```

